\documentclass[11pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[numbers,sort&compress,super]{natbib}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{MLHMM}
\usepackage{upgreek}
\usepackage{longtable}
\usepackage{array}
\usepackage[singlelinecheck=off]{caption}
\graphicspath{{SI_figures/}}

%% these lines makes it possible to cross-reference items in the main
%% text by commands like \ref{MT:label}
%\usepackage{xr}
%\externaldocument[MT:]{Manuscript}


%shorthand notation for referencing etc
\newcommand{\dc}{$~\upmu$m$^2$~s$^{-1} $}
\newcommand{\eq}[1]{Eq.~\eqref{#1}}
\newcommand{\fig}[1]{Fig.~\ref{Fig:#1}}
\newcommand*{\subfig}[2]{Fig.~\ref{Fig:#1}(#2)}
\newcommand{\FP}[1]{\textcolor{blue}{#1}}
\newcommand\ST{\rule[-0.75em]{0pt}{2.0em}} % a strut to uncrowd some tables

\newcommand{\FPR}[1]{#1}
\newcommand{\ML}[1]{#1}
\newcommand{\MLcolorOn}{}
\newcommand{\MLcolorOff}{}
% numbering w prefix
%\renewcommand{\thesection}{S\arabic{section}}  
%\renewcommand{\thetable}{S\arabic{table}}  
%\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}

\title{vbSPT (Variational Bayes for Single Particle Tracking) -- User-guide and Documentation}

\author{Fredrik Persson\footnote{freddie.persson@gmail.com ; bmelinden@gmail.com},
Martin Linden$^*$, Cecilia Unoson, and Johan Elf}
\date{\today}

\begin{document}

\maketitle 

%\section{Methods}
%\section*{About the software}
%vbSPT software is distributed under the GPL license \ldots
\noindent vbSPT is a program suite for analysis of single particle diffusion
trajectories, where the diffusion constants switch randomly according
to a discrete Markov process. The program runs on Matlab, but uses
compiled C-code to speed up the most computer intensive loops.  The
latest version of the software as well as a forum for discussion and
questions can be found at \verb+sourceforge.net/projects/vbspt/+.\medskip\\

\noindent Copyright \copyright \  2013 Martin Lind\'en, Fredrik Persson, and
Johan Elf.\medskip\\
 
\noindent This program is free software: you can redistribute it
and/or modify it under the terms of the GNU General Public License as
published by the Free Software Foundation, either version 3 of the
License, or any later version.  This program is distributed in the
hope that it will be useful, but WITHOUT ANY WARRANTY; without even
the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
PURPOSE. See the GNU General Public License for more details. You
should have received a copy of the GNU General Public License along
with this program. If not, see http://www.gnu.org/licenses/.\medskip\\

\noindent If you use it for academic research, please cite the
original article in your work:\medskip\\
\noindent Fredrik Persson, Martin Lind\'en, Cecilia Unoson, and Johan
Elf, \emph{Extracting intracellular diffusive states and transition
  rates from single molecule tracking data}, Nature Methods,
2013. doi:10.1038/nmeth.2367: \medskip\\


\noindent  This product includes software developed and copyrighted by Martin
Lind\'en (see copyright in individual files). 



\tableofcontents
\newpage
\section{vbSPT software user manual}
\label{Sec:userManual}
\input{vbSPT_userManual.tex}
\newpage
\section{Tables of variables}
\input{vbSPT_tables.tex}
\newpage
\section{Introduction to the variational algorithm for single particle tracking}
\label{Sec:statistics}
\input{vbSPT_statistics.tex}
%\newpage
\section{Derivation of the variational algorithm}
\label{Sec:derivation}
\input{vbSPT_derivation.tex}

%\section{Controls}
%\input{SI_controls.tex}
% 
%\subsection{Sample preparation}
%\input{SI_samplePrep.tex}




\newpage



\bibliographystyle{unsrtnat} \bibliography{VB-HMM}
\end{document}
